---
title: "Missing txps"
author: "Noor Pratap Singh"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/fs/cbcb-lab/rob/students/noor/Uncertainity/treeTerminusScripts')
```

#### Loading Packages
```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
suppressPackageStartupMessages(source("tree_term_climb.R"))
load("environment/brain_sim/simulate.rda")
load("environment/brain_sim/yAll.RData")
load("environment/brain_sim/y.RData")
load("environment/brain_sim/tree.RData")
load("environment/brain_sim/tAfterqval.RData")
load("environment/brain_sim/tAfterqvalNoSign.RData")
load("environment/brain_sim/tAfterqvalNoStop.RData")
levels <- node.depth(tree, 2)
levels <- ifelse(levels > 4, 5, levels)
```

#### Level distribution
```{r}
print(sapply(tAfterqval, function(node) table(levels[node[["candNodeO"]]])))
print(sapply(tAfterqvalNoSign, function(node) table(levels[node[["candNodeO"]]])))
print(sapply(tAfterqvalNoStop, function(node) table(levels[node[["candNodeO"]]])))
```

#### NTxps
```{r}
dtus <- intersect(names(iso.dtu)[iso.dtu], rownames(y))
dtuInds <- match(dtus, rownames(y))
print(sapply(c(0.01, 0.05, 0.1), function(fdr) sum(mcols(y)[["qvalue"]] <= fdr)))
print(sapply(tAfterqval, function(node) length(unlist(Descendants(tree, node[["candNodeO"]])))))
print(sapply(tAfterqvalNoSign, function(node) length(unlist(Descendants(tree, node[["candNodeO"]])))))
print(sapply(tAfterqvalNoStop, function(node) length(unlist(Descendants(tree, node[["candNodeO"]])))))
```
While for 0.01, we have more transcripts given by our method but the once fdr increases the txps given by our method begin to decrease

```{r}
alphas <- c(0.01, 0.05, 0.1)
missing <- lapply(seq_along(tAfterqvalNoStop), function(i) setdiff(which(mcols(y)[["qvalue"]] <= alphas[i]), unlist(Descendants(tree, tAfterqvalNoStop[[i]][["candNodeO"]]))))
print(sapply(missing, length))
sapply(missing, function(x) mean(x %in% dtuInds))
```
Approximately 67 and 49% that are missing from swish run for 0.05 and 0.1 qvalue cutoff aare are actually differentially expressed transcripts

#### Looking at 0.05 region
```{r}
plot(mcols(y)[["log2FC"]][missing[[2]]], -log10(mcols(y)[["qvalue"]][missing[[2]]]))
print(min(mcols(y)[["qvalue"]][missing[[2]]]))
print(max(mcols(y)[["qvalue"]][missing[[2]]]))

print(mean(tAfterqvalNoStop[[2]][["resDf"]][match(missing[[2]], tAfterqvalNoStop[[2]][["resDf"]][["inds"]]),2] < 0.1))
print(mean(tAfterqvalNoStop[[2]][["resDf"]][match(missing[[2]], tAfterqvalNoStop[[2]][["resDf"]][["inds"]]),2] < 0.07))
print(max(tAfterqvalNoStop[[2]][["resDf"]][match(missing[[2]], tAfterqvalNoStop[[2]][["resDf"]][["inds"]]),2]))
```
the minimum qvalue in the actual swish fit on the txps is 0.016, all the txps dont have that low qvalues
Looking these transcripts in the new qvalue fit we see that 63% have qvalue < 0.07 and 97% have qvalue < 0.1, with the max adjusted pvalue being 0.21