---
title: "Read Trees"
author: "Noor Pratap Singh"
date: "8/18/2021"
output: html_document
---

This file reads the tree terminus output of running on Chimpanzee bulk data and preparing the bulk and tree output accordingly
Using output from read_sal, where we have seen the existence of batch effects and we also removed medial dorsal nucleus of thalmus samples for the analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/fs/cbcb-lab/rob/students/noor/Uncertainity/treeTerminusScripts')
```

#### Reading the nwk tree
```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
trees <- read.tree("/fs/cbcb-lab/rob/students/noor/Uncertainity/brain_sim/out_term/cluster_nwk.txt")
tLengths <- sapply(trees, function(x) length(x$tip.label))
print(length(tLengths))
print(max(tLengths))
print(sum(tLengths >= 10))
print(sum(sapply(trees, function(t) length(t$tip))))
```
There are a total of 14810 trees, with 4419 groups containing atleast 10 txps and a group contains 364 txps. These groups span 55K txps
The final tree contains 66K txps

#### Processing the data running swish
```{r}
load("environment/brain_sim/y.RData")
load("environment/brain_sim/tree.RData")
yAll <- prepSwish(tree, y)
yAll <- scaleInfReps(yAll, lengthCorrect = T)
yAll <- labelKeep(yAll)
mcols(yAll)$keep=T
yAll <- swish(yAll, x = "condition")
yAll <- computeInfRV(yAll)
hist(mcols(yAll)[["pvalue"]])
save(yAll, file = "environment/brain_sim/yAll.RData")
```