---
title: "BOUTH on SWIM Data"
output: html_notebook
---

#### Loading data
```{r}
suppressPackageStartupMessages(library(BOUTH))
suppressPackageStartupMessages(source("tree_helper_function.R"))
## Reading data
quantDir <- "../boot_gibbs/quant_output/human/swim_gencodev28"
samples <- as.vector(outer(c(1:6), c("01","02"), function(x,y) paste("out",x,"sample",y,sep="_")))
files <- file.path(quantDir, samples, "quant.sf")
colData <- data.frame(files = files, names = samples, condition = as.factor(rep(c(1,2),each=6)))
# seSwim <- tximeta::tximeta(colData)
# save(seSwim, file = "environment/se_swim.RData")
load("environment/se_swim.RData")
```

#### Reading trees
Merging all the individual trees by treating them as children of one parent root node
```{r}
trees <- read.tree("../terminus/data/term/cluster_nwk.txt")
#tree <- mergeTree(trees)
tLengths <- sapply(trees, function(x) length(x$tip.label))
print(hist(tLengths))
```

#### Running Swish
Swish involves running on filtered txps but filtering removes txps that are in the tree
I thus first filter using the swish criteria and then further add the txps that are in the tree but missing in the filtered SE
Swish is then run on the new set
```{r}
# y <- scaleInfReps(seSwim)
# y <- labelKeep(y)
# y <- runSwishtree(tree, y, type = "union")
# save(y, file = "environment/swish_y.RData")
load("environment/swish_y.RData")
```

#### Creating final merged tree
Each txp for which the pvalue is computed but is not in the tree is also added as a leaf
```{r}
# length(setdiff(as.numeric(tree$tip.label), match(rownames(y), rownames(seSwim)))) == 0
# tree <- mergeLeaves(tree, y, seSwim)
# save(tree, file = "environment/tree.RData")
load("environment/tree.RData")
```

#### Creating data frame needed by BOUTH
```{r}
# treeDf <- getTreeDf(tree)
# save(treeDf, file="environment/treeDf.RData")
load("environment/treeDf.RData")
print(head(treeDf))
print(dim(treeDf))
```

#### Running BOUTH yields only 1 driver node aka root
```{r}
pvalues <- mcols(y)[["pvalue"]][as.numeric(rownames(treeDf))]
plot(hist(pvalues))

# bouth_test <- list()
# bouth_test[['fdr_0.1']] <- bouth(anno.table = treeDf, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.1, is.weighted = TRUE)
# bouth_test[['fdr_0.05']] <- bouth(anno.table = treeDf, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.05, is.weighted = TRUE)
# save(bouth_test, file="environment/bouth_test_all.RData")
load("environment/bouth_test_all.RData")
print(head(bouth_test$fdr_0.1$results.by.node))
print(head(bouth_test$fdr_0.1$results.by.level))
print(sum(bouth_test$fdr_0.1$results.by.node[,"is.driver"]))
print(hist(bouth_test$fdr_0.1$results.by.node[,"pvalue"]))
```

#### Filtered set
Since many leaves existed that had directly the root as the parent, I only focus on the txps in tree and run swish on only them
```{r}
# tree <- mergeTree(trees)
# y <- scaleInfReps(seSwim)
# y <- labelKeep(y)
# y <- runSwishtree(tree, y, type = "tree")
# tree <- convTree(tree, y, seSwim) ##Changing the indexes that are reflected
# treeDfGroup <- getTreeDf(tree)
# save(tree, file="environment/tree_group.RData")
# save(treeDfGroup, file="environment/treeDfGroup.RData")
# save(y, file="environment/ygroup.RData")
load("environment/tree_group.RData")
load("environment/treeDfGroup.RData")
load("environment/ygroup.RData")
print(dim(treeDfGroup))

pvalues <- mcols(y)[["pvalue"]][as.numeric(rownames(treeDfGroup))]
plot(hist(pvalues))
# bouth_test <- list()
# bouth_test[['fdr_0.1']] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.1, is.weighted = TRUE)
# bouth_test[['fdr_0.01']] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.01, is.weighted = TRUE)
# bouth_test[['fdr_0.05']] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.05, is.weighted = TRUE)
# bouth_test <- bouth_test[c(2,3,1)]
# save(bouth_test, file = "environment/bouth_test_group.RData")
load("environment/bouth_test_group.RData")
print(sum(bouth_test$fdr_0.01$results.by.node[,"is.driver"]))
print(hist(pvalues))
print(hist(bouth_test$fdr_0.01$results.by.node[,"pvalue"]))
```
Even then we get 1 driver node

#### Using DESeq2 instead of swish
Since many txps get pvalue as NA, they are set to 1
```{r}
# txi <- tximeta::tximeta(colData, type = "salmon", txOut = TRUE, countsFromAbundance="lengthScaledTPM")
# ddsSim <- DESeqDataSet(txi[rownames(y),], design=~condition)
# ddsSim <- DESeq(ddsSim)
# res <- results(ddsSim, name="condition_2_vs_1")
# res$pvalue[is.na(res$pvalue)] <- 1
# res$padj[is.na(res$padj)] <- 1
# save(res, file="environment/resDeseq2_sim.RData")
load("environment/resDeseq2_sim.RData")
pvalues <- res[rownames(y), "pvalue"]
bouth_sim <- list()
# bouth_sim[["0.1"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.1, is.weighted = TRUE)
# bouth_sim[["0.05"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.05, is.weighted = TRUE)
# bouth_sim[["0.01"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.01, is.weighted = TRUE)
# bouth_sim <- bouth_sim[c(2,3,1)]
# save(bouth_sim, file="environment/bouth_sim.RData")
tseSim <- createTreeSummarizedObject(y, treeDfGroup, tree)
load("environment/bouth_sim.RData")
print(sum(bouth_sim$results.by.node[,"is.driver"]))
print(hist(pvalues))
print(hist(bouth_sim$results.by.node[,"pvalue"]))
```

#### On the true data
Issues - Missing txps since only 47K have non zero counts, the remaining are added and set to 0 (around 7k out of 23K)
No length normalization
```{r}
dir <- "/fs/cbcb-lab/rob/students/noor/Uncertainity/swim_data"
samples <- as.vector(outer(c(1:6), c("1","2"), function(x,y) paste(paste("out",x, sep = "_"), paste(paste("out",y, sep=""), "counts.tsv", sep = "_"), sep="/")))
files <- file.path(dir, samples)
dd <- readTrueCounts(files, samples)
mTxps <- setdiff(rownames(y),rownames(dd)) ## txps that are missing
print(length(mTxps))
# mDf <- matrix(0,nrow=length(mTxps),ncol=ncol(dd), dimnames=list(mTxps, colnames(dd)))
# dd <- rbind(dd, mDf)
# ddsTrue <- DESeqDataSetFromMatrix(countData = dd[rownames(y),], colData = colData, design=~condition)
# ddsTrue <- DESeq(ddsTrue)
# resTrue <- results(ddsTrue)
# resTrue$pvalue[is.na(resTrue$pvalue)] <- 1
# resTrue$padj[is.na(resTrue$padj)] <- 1
# save(resTrue, file="environment/resDeseq2_true.RData")
load("environment/resDeseq2_true.RData")
pvalues <- resTrue[rownames(y),"pvalue"]
bouth_true <- list()
# bouth_true[["0.01"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.01, is.weighted = TRUE)
# bouth_true[["0.05"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.05, is.weighted = TRUE)
# bouth_true[["0.1"]] <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.1, is.weighted = TRUE)
# bouth_true <- bouth_true[c(2,3,1)]
# save(bouth_true, file="environment/bouth_true.RData")
load("environment/bouth_true.RData")
print(sum(bouth_true$results.by.node[,"is.driver"]))
print(hist(pvalues))
print(hist(bouth_true$results.by.node[,"pvalue"]))
```