---
title: "Comparison"
output: html_document
---

```{r}
library(BOUTH)
source("tree_helper_function.R")
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/se_swim.RData")
load("environment/ygroup.RData")
trees <- read.tree("../terminus/data/term/cluster_nwk.txt")
tLengths <- sapply(trees, function(x) length(x$tip.label))
tree <- mergeTree(trees, se = seSwim)

```

### Loading BOUTH run using Swish and DESeq2
```{r}
load("environment/bouth_sim.RData") ## DESeq2
load("environment/bouth_test_group.RData") ## Swish
bouth_swish <- bouth_test$fdr_0.05
bouth_deseq <- bouth_sim
```

### Overlap between detected nodes
```{r}
print(head(bouth_swish$results.by.level))
print(head(bouth_deseq$results.by.level))

print(tail(bouth_swish$results.by.level))
print(tail(bouth_deseq$results.by.level))

dediff <- which(bouth_deseq$tree@test$reject)
swdiff <- which(bouth_swish$tree@test$reject)
print(paste(length(dediff), length(swdiff), length(intersect(swdiff, dediff))))
```
There exists a huge overlap between both the methods with 2900 nodes overlapping


```{r}
.StoufferCombination<-function(pvalue){
  nan.pvalue = pvalue[!is.na(pvalue)]
  n = sum(nan.pvalue<=1)
  if(n>0){
    z.stat = sum(qnorm(1-nan.pvalue))/sqrt(n)
    return(1-pnorm(z.stat))
  }else{
    return(0)
  }
}
```

```{r}
pdeseq <- bouth_sim$tree@test$pvalue
pswish <- bouth_swish$tree@test$pvalue
print(hist(pdeseq))
print(hist(pswish))
plot(pdeseq, pswish)

lev_inds <- which(bouth_sim$tree@node$level==92) ##Indices just above root

lev_diff_deseq <- intersect(lev_inds, dediff)
lev_diff_swish <- intersect(lev_inds, swdiff)

lev_undiff_deseq <- setdiff(lev_inds, lev_diff_deseq)
lev_undiff_swish <- setdiff(lev_inds, lev_diff_swish)
```

### Pvalues for detected nodes
```{r}
print(hist(pdeseq[lev_diff_deseq]))
print(hist(pswish[lev_diff_swish]))
```

### Pvalues for undetected nodes
```{r}
print(hist(pdeseq[lev_undiff_deseq]))
print(hist(pswish[lev_undiff_swish]))

alphaDeseq <- bouth_deseq$results.by.level[92,4]
alphaSwish <- bouth_swish$results.by.level[92,4]
print(paste(alphaDeseq, alphaSwish))

pDeUniform <- (pdeseq[lev_undiff_deseq] - alphaDeseq)/(1-alphaDeseq)
pSwUniform <- (pswish[lev_undiff_swish] - alphaSwish)/(1-alphaSwish)

print(paste(.StoufferCombination(pSwUniform), .StoufferCombination(pDeUniform), sep = ","))
print(hist(pDeUniform))
print(hist(pSwUniform))
```
cutoffs at the node below are similar but pvalue for the root is very different, similarly the cutoff for the root is very different.
Problem is both the pvalue and the threshold

```{r}
#b <- bouth(anno.table = treeDfGroup, pvalue.leaves = pvalues, na.symbol = "Undefined", far = 0.05, is.weighted = FALSE)
#pvalues <- mcols(y)[rownames(treeUDf),"pvalue"]
#save(p, file="environment/b_unweighted.RData")
load("environment/b_unweighted.RData")
print(tail(b$results.by.level))
```
