```{r}
source("tree_helper_function.R")
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/whole_txp/swish_y.RData")
load("environment/whole_txp/tree.RData")
library(ggplot2)
library(dplyr)

missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(y)+1:tree$Nnode
aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1]))
fold_changes <- fold_changes[rownames(y),]

levels <- node.depth(tree, 2)
levels <- sapply(levels, function(x) {
    if(x <= 4)
        as.character(x)
    else
        "5"
})
levInds <- lapply(unique(levels), function(lev) which(levels == lev))
fcDf <- data.frame(LFC = logFCNodes, levels = levels)
```

#### Plots for log fold change at different levels
```{r}
ggplot(fcDf, aes(LFC)) +
        geom_histogram() +
    facet_wrap(~levels)

ggplot(fcDf, aes(levels, LFC)) +
        geom_violin()

ggplot(fcDf, aes(levels, LFC)) +
        geom_boxplot()


fcDf %>% group_by(levels) %>% 
    summarize(prop_FCs = mean(LFC < 0.067))
```
At each level, we will have 10-12% differentially expressed nodes. We can decrease it more by further decreasing logFC criteria, since we wont be testing root anyway if 10-12% nodes is big

#### Plotting pvalues against infRVs and levels
```{r}
load("environment/whole_txp/ySwishAllNodes.RData")
df <- data.frame(IRVCut = cut(mcols(yAll)[["meanInfRV"]], breaks = quantile(mcols(yAll)[["meanInfRV"]], 0:4/4), include.lowest = T),
                 levels, pvalue = mcols(yAll)[["pvalue"]], infRVs = mcols(yAll)[["meanInfRV"]])

##pvalue histograms
ggplot(df, aes(pvalue)) +
  geom_histogram() + 
  facet_grid(IRVCut ~ levels)

## Only infRV
ggplot(df, aes(pvalue)) +
  geom_histogram() + 
  facet_wrap(~IRVCut)

### Violin plots
ggplot(df, aes(IRVCut, -log10(pvalue))) +
  geom_violin() + facet_wrap(~levels)

ggplot(df, aes(IRVCut, -log10(pvalue))) +
  geom_violin()

df %>% group_by(IRVCut, levels) %>%
  summarize(mean_prop_alt = mean(pvalue < 0.001)) %>%
  ggplot(aes(IRVCut, levels, fill=mean_prop_alt)) +
  geom_tile()

df %>% group_by(IRVCut) %>%
  summarize(mean_prop_alt = mean(pvalue < 0.001)) %>%
  ggplot(aes(IRVCut, mean_prop_alt)) +
  geom_bar(stat = "identity")

df <- df %>% mutate(group=factor(paste(as.integer(IRVCut), as.integer(levels), sep="-")))
table(df$group)
```
Irresepective of level, the peak of low pvalues seem to fall as infRV increases (histogram and violin). Highest infRV has the least power. From Violin plot, it seems based on levels, sometimes middle infRVs might have higher peaks.

As levels increase, across most infRV bins power seems to increase. Across infRV for most levels, power increases 

#### Running IHW
```{r}
library(IHW)
### IHW only stratified by grouping infRV and levels
res <- ihw(pvalue ~ group,  data = df,
           alpha = 0.1,
           covariate_type = "nominal")
print(rejections(res))
padjBH <- p.adjust(df$pvalue, method = "BH")
print(sum(padjBH <= 0.1))
plot(res)

### IHW only stratified by infRV
res2 <- ihw(pvalue ~ infRVs, data = df,
           alpha = 0.1,
           covariate_type = "ordinal")
print(rejections(res2))
plot(res2)
```
Power seems to increase for IHW compared to BH, many have weight 1

#### Plottting IHW
```{r}
ihw_dat <- as.data.frame(res)
ihw_dat$cov1 <- df$IRVCut
ihw_dat$cov2 <- as.factor(df$levels)
ihw_dat %>% group_by(group) %>%
  summarize(mean_wt = mean(weight), mean_small=mean(pvalue < 1e-4),
            cov1=first(cov1), cov2=first(cov2)) %>%
  ggplot(aes(mean_small, mean_wt, color=cov1, shape=cov2)) + geom_point()

ihw_dat2 <- as.data.frame(res2)
ihw_dat2$cov1 <- df$IRVCut
ihw_dat2 %>% group_by(group) %>%
  summarize(mean_wt = mean(weight), mean_small=mean(pvalue < 1e-4),
            cov1=first(cov1)) %>%
  ggplot(aes(mean_small, mean_wt, color=cov1)) + geom_point()
```

#### Preparing folds for IHW by getting them on same branches
```{r}
df[["IRVCut"]] = cut(mcols(yAll)[["meanInfRV"]], breaks = quantile(mcols(yAll)[["meanInfRV"]], 0:8/8))
rChildNodes <- Descendants(tree, nrow(y)+1, "children") ### Root child nodes
folds <- rep(1, nrow(yAll))
rParNodes <- rChildNodes[rChildNodes %in% seq(nrow(y))] ## Txps that directly map to root
remChildNodes <- setdiff(rChildNodes, rParNodes)
ch <- cut(remChildNodes, breaks = quantile(remChildNodes, 0:4/4))
d <- split(remChildNodes, ch)
for(i in seq_along(d))
  folds[c(d[[i]],unlist(Descendants(tree, d[[i]],"all")))] <- i+1

print(table(folds))
resgroup <- ihw(pvalue ~ infRVs, data = df,
           alpha = 0.1, folds = folds,
           covariate_type = "ordinal")
plot(resgroup)
print(rejections(resgroup))
ihw_datgroup <- as.data.frame(resgroup)
ihw_datgroup$cov1 <- df$IRVCut
ihw_datgroup %>% group_by(group) %>%
  summarize(mean_wt = mean(weight), mean_small=mean(pvalue < 1e-4),
            cov1=first(cov1)) %>%
  ggplot(aes(mean_small, mean_wt, color=cov1)) + geom_point()

print(sum(ihw_datgroup$adj_pvalue<=0.1 & padjBH<=0.1))
```
Net total increase in 1, with a net overlap of 11335 nodes

