---
title: "Inner Nodes"
output: html_notebook
---

```{r}
source("tree_helper_function.R")
source("bouth_helper.R")
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/se_swim.RData")
load("environment/ygroup.RData")
load("environment/bouth_true.RData")
load("environment/bouth_sim.RData")
load("environment/treeDfGroup.RData")
load("environment/tree_group.RData")
library(ggplot2)
```


#### Modifying the true counts matrices
```{r}
missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
print(length(missingTxps))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(y)+1:tree$Nnode
aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1])) 
fold_changes <- fold_changes[rownames(y),]

trueCountNodes <- which(abs(logFCNodes) >= 0.066) ##Making sure root is non differentiated
trueLeaves <- union(trueCountNodes[trueCountNodes <= nrow(y)],unlist(Descendants(tree,trueCountNodes[trueCountNodes > nrow(y)])))

trueDetNodes <- sapply(bouth_true, function(b) extractBouthNodes(b, y))
simDetNodes <- sapply(bouth_sim, function(b) extractBouthNodes(b, y))

aggSimCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), assays(y)[["counts"]], list(c(1:6),c(7:12)))
logFCSimNodes <- ifelse(rowSums(aggSimCountsNodes)==0, 0, log2(aggSimCountsNodes[,2]/aggSimCountsNodes[,1])) 

###Used in swish paper 
# iso.any <- iso.dtu | iso.dte | iso.dge
```

### Exploring the inner driver nodes


#### Looking at the distribution within nodes
```{r}
driverInn <- simDetNodes[["0.05"]][simDetNodes[["0.05"]] %in% innNodes] ##Inner driver nodes
leaves <- Descendants(tree, driverInn, "tips")
dLengths <- sapply(leaves, length)
hist(dLengths)

detNodes <- extractBouthNodes(bouth_sim[["0.05"]],se,"detected") ##Detected nodes
lNotDeg <- list()
for(i in seq_along(leaves)) {
    lNotDeg[[i]] <- setdiff(leaves[[i]], detNodes)
}

nInds <- which(sapply(lNotDeg, length) != 0)
print(hist(sapply(lNotDeg,length)[nInds]))
print(length(nInds))

b <- bouth_sim[["0.05"]]
notDetNodes <- b$results.by.node[match(rownames(y)[unlist(lNotDeg)], b$results.by.node$LEAF),]
plot(log10(notDetNodes[["pvalue"]]), logFCSimNodes[unlist(lNotDeg)]) + abline(v=log10(0.05))
print(mean(abs(logFCSimNodes[unlist(lNotDeg)]) > 1))
```
76 driver nodes that have atleast 1 leaf that is not differentially expressed. Further, while most of these nodes have 1 txp as non-deg, some have as many as 7 such txps. Around 70% leaves that have not been detected have logFC larger than 1

#### Exploratory analysis on the type of inner nodes
```{r}
tNodes <-  Descendants(tree, nrow(y)+1, type="c")##Children of root aka all the indivudal tree groups
diffTNodes <- intersect(tNodes, driverInn) ## Tree groups that are differentially expressed
print(length(diffTNodes))
othDiffNodes <- setdiff(driverInn, diffTNodes) ## Differentially expressed inner nodes, that are not roots of individual tree groups
print(length(othDiffNodes))

hist(dLengths[match(diffTNodes,driverInn)]) ## Leave nodes for the actual root
hist(dLengths[match(othDiffNodes,driverInn)]) ## Leave nodes for the inner nodes

distRoot <- sapply(Ancestors(tree, othDiffNodes), length) - 1 ## Distance from the root of the tree group for each indivudal node
hist(distRoot)
```
Of the 461 differentially expressed inner nodes, 263 nodes belong to the tree groups, 198 nodes are broken earlier. Of these 198 nodes while most are just located at a distance 1, many are located at a distance of 2 or greater than 2

#### Change in TPR,FPR as we go one level above or below
```{r}
### Going one level below
dn <- Descendants(tree, driverInn, type = "children")
lowerLevel <- sapply(dn, function(x) ifelse(length(x) > 1, sample(x,1), x))
sNodes <- union(intersect(simDetNodes[["0.05"]], seq(nrow(y))), lowerLevel)
print(unlist(computeTPFP(trueDetNodes[["0.05"]], sNodes, y, logFCNodes, tree, type = "leaves"))) ### At leaf level
print(unlist(computeTPFP(trueDetNodes[["0.05"]], sNodes, y, logFCNodes, tree))) ### At node level
print(unlist(computeTPFP(trueCountNodes, sNodes, y, logFCNodes, tree, type = "leaves")))

### Going one level above
upLevel <- union(diffTNodes,Ancestors(tree, othDiffNodes, type = "parent")) 
print(length(upLevel)) ##Number of nodes decreases
sNodes <- union(upLevel, intersect(simDetNodes[["0.05"]], seq(nrow(y))))
print(unlist(computeTPFP(trueDetNodes[["0.05"]], sNodes, y, logFCNodes, tree, type = "leaves")))
print(unlist(computeTPFP(trueDetNodes[["0.05"]], sNodes, y, logFCNodes, tree)))
print(unlist(computeTPFP(trueCountNodes, sNodes, y, logFCNodes, tree, type = "leaves")))
```
We want to check this but ideally the method itself should take care of it. At the node level, FDR increases and TPR decreases either going up or down. Looking at the leaves, at the lower level while the FDR decreases, there is a considerable decrease in TPR as well. Similarly, at increasing the level, both TPR and FDR increase with the latter being much more significant.

When comparing with the **actual truth**, I observe a both decrease in TPR and increase in FDR at a lower level
However on **increasing** the level, while both FDR and TPR increase, the increase in latter is significant

#### Looking at the nodes that have txps with varying signs of effect size
```{r}
lInds <- c()
for(i in seq_along(leaves)) {
    lp <- 0
    ln <- 0
    
    if(sum(logFCSimNodes[leaves[[i]]] > 0))
        lp <- 1
    if(sum(logFCSimNodes[leaves[[i]]] < 0))
        ln <- -1
    if(lp*ln == -1)
        lInds <- c(lInds, driverInn[i])
}
hist(logFCSimNodes[driverInn]) ## All Inner Nodes
hist(logFCSimNodes[lInds]) ## All inner nodes that have atleast txps with varying signs of logFCs
print(mean(abs(logFCSimNodes[driverInn]) <= 0.5))
print(mean(abs(logFCSimNodes[lInds]) <= 0.5))

```
Most of the driver nodes have only two underlying txps. Further, around 24% nodes have logFC < 0.5 which increases to 0.73 for nodes containing txps with varying signs. 

#### Looking at these nodes further sortring by inFRV
```{r}
mInds <- match(lInds, driverInn)
y <- computeInfRV(y)
infs <- sapply(mInds, function(x) sum(mcols(y)[leaves[[x]],"meanInfRV"]))
infInds <- order(infs, decreasing = T)
head(infs[infInds])
tail(infs[infInds])

i=1
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="((((((1,2),3),4),5),(((6,7),8),9)),(10,11));")
plot(t)

i=2
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="(1,(2,3));") ###Not connected to root
plot(t)

i=3
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])

i=5
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="((1,2),(3,4));")
plot(t)

i=7
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])

i=8
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="((1,2),3);")
plot(t)

i=9
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="((1,2),3);")
plot(t)

i=18
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="(((1,(2,3)),(4,(5,6))),7);")
plot(t)

i=20
print(logFCSimNodes[leaves[[mInds[infInds][i]]]])
t=read.tree(text="(((1,(2,3)),4),5);")
plot(t)
```

All the nodes having the least infRV have only 2 leaves. This might me also useful with nodes having leaves with same signs.
Count based aggregation might make us lose many such nodes where children have significant logFCs with opposite signs. Do we still consider these as driver nodes? Further, nodes with very opposing fold changes getting merged is sligthly conufsing
