---
title: "Testing BOUTH Differential"
author: "Noor Pratap Singh"
date: "5/29/2021"
output: html_document
---

```{r}
source("tree_helper_function.R")
source("bouth_helper.R")
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/se_swim.RData")
load("environment/ygroup.RData")
load("environment/bouth_true.RData")
load("environment/bouth_sim.RData")
load("environment/treeDfGroup.RData")
load("environment/tree_group.RData")
library(ggplot2)
```


#### Modifying the true counts matrices
```{r}
missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
print(length(missingTxps))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(sim.counts.mat)+1:tree$Nnode
aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1])) 
fold_changes <- fold_changes[rownames(y),]

trueCountNodes <- which(abs(logFCNodes) >= 0.066) ##Making sure root is non differentiated
print(length(trueCountNodes)) ## Total differential nodes
print(sum(trueCountNodes <= nrow(y))) ## Differential txps
trueLeaves <- union(trueCountNodes[trueCountNodes <= nrow(y)],unlist(Descendants(tree,trueCountNodes[trueCountNodes > nrow(y)])))
print(length(trueLeaves))

###Used in swish paper 
# iso.any <- iso.dtu | iso.dte | iso.dge
```
Around 8K txps are missing, we account for these by adding 0s, setting logFC also equal to 0 
A total of 38K nodes with around 23K txps
Around 7K nodes are differentially expressed containing 3300 txps and if the all the underlying leaves are considered then 6400 txps

### Evaluations
#### Driver nodes, with ground truth being bouth run on true data
Here we compare when running BOUTH on pvalues obtained by running DESeq2 on true data as our ground truth
We have further two comparisons
    - comparing at the node level (where driver nodes aka both inner nodes/leaves are directly compared)
    - comparing the leaves (where an inner node is replaced by the descendant leaf). We will call this **leaf** evaluation henceforth.
```{r}
trueDetNodes <- sapply(bouth_true, function(b) extractBouthNodes(b, y))
simDetNodes <- sapply(bouth_sim, function(b) extractBouthNodes(b, y))

## Using only driver Nodes
stats <- mapply(function(x,z) {
    computeTPFP(x, z, y, logFCNodes, tree, pTab=T)
}, trueDetNodes, simDetNodes)
print(stats)

## Using leaves of driver nodes
stats <- mapply(function(x,z) {
    computeTPFP(x, z, y, logFCNodes, tree, type="leaves", pTab=T)
}, trueDetNodes, simDetNodes)
print(stats)
```

#### Truth being defined on true data and using leaves
Here the ground truth is the actual true Nodes, where a node(inner or leaf) is called differentially expressed when the logFC is > 0.066. We compare with the ground truth both the bouth results on the simulated and true data
```{r}
### Simulated data
stats <- sapply(simDetNodes, function(x) computeTPFP(trueCountNodes, x, y, logFCNodes, tree, type = "leaves"))
print(stats)
### True data
stats <- sapply(trueDetNodes, function(x) computeTPFP(trueCountNodes, x, y, logFCNodes, tree, type = "leaves"))
print(stats)
```


#### logFC differences
```{r}
aggSimCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), assays(y)[["counts"]], list(c(1:6),c(7:12)))
logFCSimNodes <- ifelse(rowSums(aggSimCountsNodes)==0, 0, log2(aggSimCountsNodes[,2]/aggSimCountsNodes[,1])) 
```
##### All nodes
```{r}
print(plot(logFCNodes, logFCSimNodes) + abline(0,1))
```

##### Signifcant nodes by true counts
```{r}
print(plot(logFCNodes[trueCountNodes], logFCSimNodes[trueCountNodes]) + abline(0,1))
```

##### Signifcant nodes that are detected by BOUTH
```{r}
print(plot(logFCNodes[simDetNodes[["0.05"]]], logFCSimNodes[simDetNodes[["0.05"]]]) + abline(0,1))
```

##### Non-Signficant nodes
```{r}
s <- setdiff(seq(length(logFCNodes)), trueCountNodes)
print(plot(logFCNodes[s], logFCSimNodes[s]) + abline(0,1))
```

#### Differences between the log fold changes
Some differences go to infinity as well
```{r}
diff <- logFCSimNodes-logFCNodes
diff[is.nan(diff)] <- 0
ggplot(data.frame(x=rep("x", length(diff)), diff = diff), aes(x = x,y=diff)) + geom_violin()
ggplot(data.frame(x=rep("x", length(diff)), diff = diff), aes(x = x,y=diff)) + geom_boxplot()
```


### Breaking down things by logFC
Using the indices obtained on the observed data and stratifying by the logFC changes

#### Sim vs True all nodes

##### Node Level
```{r}
cutL <- cut_number(abs(logFCSimNodes), 4)
print(table(cutL))
lInds <- sapply(levels(cutL), function(l) which(cutL==l))
nRes <- vector(mode = "list", length = length(simDetNodes))
names(nRes) <- names(simDetNodes)
for(i in seq_along(simDetNodes))
{
    sim <- simDetNodes[[i]]
    true <- trueDetNodes[[i]]
    nRes[[i]] <- sapply(lInds, function(x) 
    {
        sim <- intersect(x, simDetNodes[[i]])
        true <- intersect(x, trueDetNodes[[i]])
        computeTPFP(true, sim, y, logFCNodes, tree)
    })
}
print(nRes)
```

##### Leaf Level
```{r}
nResL <- vector(mode = "list", length = length(simDetNodes))
names(nResL) <- names(simDetNodes)
for(i in seq_along(simDetNodes))
{
    sim <- simDetNodes[[i]]
    true <- trueDetNodes[[i]]
    nResL[[i]] <- sapply(lInds, function(x) 
    {
        sim <- intersect(x, simDetNodes[[i]])
        true <- intersect(x, trueDetNodes[[i]])
        computeTPFP(true, sim, y, logFCNodes, tree, type = "leaves")
    })
}
print(nResL)
```

##### Simulated and true counts
```{r}
nResS <- vector(mode = "list", length = length(simDetNodes))
names(nResS) <- names(simDetNodes)
for(i in seq_along(simDetNodes))
{
    sim <- simDetNodes[[i]]
    nResS[[i]] <- sapply(lInds, function(x) 
    {
        sim <- intersect(x, simDetNodes[[i]])
        true <- intersect(x, trueCountNodes)
        computeTPFP(true, sim, y, logFCNodes, tree, type = "leaves")
    })
}
print(nResS)
```

##### BOUTH on True with True Counts
```{r}
nResT <- vector(mode = "list", length = length(simDetNodes))
names(nResT) <- names(simDetNodes)
for(i in seq_along(simDetNodes))
{
    sim <- simDetNodes[[i]]
    nResT[[i]] <- sapply(lInds, function(x) 
    {
        sim <- intersect(x, trueDetNodes[[i]])
        true <- intersect(x, trueCountNodes)
        computeTPFP(true, sim, y, logFCNodes, tree, type = "leaves")
    })
}
print(nResT)
```