---
title: "treeClimbR Results"
author: "Noor Pratap Singh"
date: "6/14/2021"
output: html_document
---

```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
library(phangorn)
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/whole_txp/swish_y.RData")
load("environment/whole_txp/tree.RData")
missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(y)+1:tree$Nnode
iso.any <- iso.dtu | iso.dte | iso.dge
diff_txps <- rownames(fold_changes)[iso.any[rownames(fold_changes)]]
tLeaf <- which(rownames(y) %in% diff_txps)

aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1]))
fold_changes <- fold_changes[rownames(y),]

trueCountNodes <- which(abs(logFCNodes) >= 0.066) ##Making sure root is non differentiated
trueDriverNodes <-  findDriverGround(trueCountNodes, tree, logFCNodes)[[1]] ### All underlying children have same sign

print(length(trueDriverNodes))
print(length(tLeaf))
```

#### Extracting detected nodes
```{r}
load("environment/whole_txp/bSwish.RData")
load("environment/whole_txp/bSwishInfRV.RData") 
load("environment/whole_txp/b_deseq2_infRV.RData") 
load("environment/whole_txp/bDESeq2.RData")
load("environment/whole_txp/ySwishAllNodes.RData")
load("environment/innNodes/res.RData")

detNodes <- list()
detNodes[["OnlySwishL"]] <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(y)[,"qvalue"] <= x)) ##Leaves
detNodes[["tcrSw"]] <- sapply(bSwish, function(sw) sw$output[sw$output$signal.node, "node"]) ##Treeclimbr Swish on all nodes
detNodes[["tcrSwInf"]] <- sapply(bSwishInf, function(sw) sw$output[sw$output$signal.node, "node"]) ##Treeclimbr Swish on nodes stratified by infrv (6 bins)
detNodes[["tcrDeseq2"]] <- sapply(bDeseq2, function(de) de$output[de$output$signal.node, "node"]) ##Treeclimbr deseq2 nodes
detNodes[["tcrDeInf"]] <- sapply(bDeseq2Inf, function(de) de$output[de$output$signal.node, "node"]) ##Treeclimbr deseq2 on nodes stratified by infrv (6 bins)

# detNodesDeN <- lapply(c(0.01,0.05,0.1), function(x) which(res$padj <= x))
# detNodes[["OnlySwishN"]] <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(yAll)[,"qvalue"] <= x)) ##On nodes
```

#### Comparing with leaf nodes
```{r}
df <- data.frame(Method=character(), Metric=character(), FDR_0.01 = numeric(), FDR_0.05 = numeric(), FDR_0.10 = numeric())
for(i in seq_along(detNodes))
{
    stats <-  sapply(detNodes[[i]], function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
    colnames(stats) <- colnames(df)[c(3:5)]
    stats <- cbind(Method=rep(names(detNodes)[i],2), Metric = c("FDR", "TPR"), stats)
    df <- rbind(df, stats)
}
print(df)
### tcr treeClimbR
```

##### treeclimbR using different bins for stratifying based on infRV
```{r}
load("environment/whole_txp/b_swish_infRV_diff.RData")
bins <- as.character(c(4,6,8,12,16))
detNodesInfBins <- vector("list", length(bins))
names(detNodesInfBins) <- bins

dfInf <- data.frame(Method=character(), Metric=character(), FDR_0.01 = numeric(), FDR_0.05 = numeric(), FDR_0.10 = numeric())
for(i in seq_along(detNodesInfBins))
{
    detNodesInfBins[[bins[i]]] <- sapply(bSwishInf[[i]], function(sw) sw$output[sw$output$signal.node, "node"])
    stats <-  sapply(detNodesInfBins[[bins[i]]], function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
    colnames(stats) <- colnames(dfInf)[c(3:5)]
    stats <- cbind(Method=rep(bins[i],2), Metric = c("FDR", "TPR"), stats)
    dfInf <- rbind(dfInf, stats)
    
}
print(dfInf)
```

##### treeclimbR using levels for stratification
```{r}
load("environment/whole_txp/b_swish_levels.RData")
detLevels <- sapply(bSwishLev, function(sw) sw$output[sw$output$signal.node, "node"])
stats <-  sapply(detLevels, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

Tree level methods in general seem to have much higher FDR just swish but also a good TPR. DESeq2 has the highest TPR and at 0.1 level, FDR is also comparable to Swish. As expected cut number 4 seems optimal when chosing bins for stratification. Level based has the lowest FDR.

One side note is that when using the whole transcriptome as a reference, we are not actually making use of the tree structure completely since around 33K (57%) have root as their parent. So, do we expect tree based method to account for these as well.