---
title: "treeClimbR Results"
author: "Noor Pratap Singh"
date: "6/14/2021"
output: html_document
---

```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
library(phangorn)
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/ygroup.RData")
load("environment/tree_group.RData")
missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(y)+1:tree$Nnode

aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1]))
fold_changes <- fold_changes[rownames(y),]

trueCountNodes <- which(abs(logFCNodes) >= 0.066) ##Making sure root is non differentiated
trueDriverNodes <-  findDriverGround(trueCountNodes, tree, logFCNodes)[[1]] ### All underlying children have same sign
trueDriverNodes <- sort(trueDriverNodes)

iso.any <- iso.dtu | iso.dte | iso.dge
tLeaf <- which(rownames(y) %in% names(iso.dtu)[iso.any])
print(length(trueDriverNodes))
print(length(tLeaf))
```

#### Extracting detected nodes
```{r}
load("environment/parTClimbR/bSwish.RData")
load("environment/parTClimbR/bSwishInfRV.RData")
load("environment/parTClimbR/bDESeq2.RData")
load("environment/parTClimbR/b_deseq2_infRV.RData")
load("environment/innNodes/ySwish.RData")
load("environment/innNodes/res.RData")

detNodes <- list()
detNodes[["OnlySwishL"]] <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(y)[,"qvalue"] <= x)) ##Leaves
detNodes[["tcrSw"]] <- sapply(bSwish, function(sw) sw$output[sw$output$signal.node, "node"]) ##Treeclimbr Swish on all nodes
detNodes[["tcrSwInf"]] <- sapply(bSwishInf, function(sw) sw$output[sw$output$signal.node, "node"]) ##Treeclimbr Swish on nodes stratified by infrv
detNodes[["tcrDeseq2"]] <- sapply(bDeseq2, function(de) de$output[de$output$signal.node, "node"]) ##Treeclimbr deseq2 nodes
detNodes[["tcrDeInf"]] <- sapply(bDeseq2Inf, function(de) de$output[de$output$signal.node, "node"]) ##Treeclimbr deseq2 on nodes stratified by infrv

# detNodesDeN <- lapply(c(0.01,0.05,0.1), function(x) which(res$padj <= x))
# detNodes[["OnlySwishN"]] <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(yAll)[,"qvalue"] <= x)) ##On nodes
```

#### Comparing with leaf nodes
```{r}
df <- data.frame(Method=character(), Metric=character(), FDR_0.01 = numeric(), FDR_0.05 = numeric(), FDR_0.10 = numeric())
for(i in seq_along(detNodes))
{
    stats <-  sapply(detNodes[[i]], function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
    colnames(stats) <- colnames(df)[c(3:5)]
    stats <- cbind(Method=rep(names(detNodes)[i],2), Metric = c("FDR", "TPR"), stats)
    df <- rbind(df, stats)
}
print(df)
### tcr treeClimbR
```

##### treeclimbR using different bins for stratifying based on infRV
```{r}
load("environment/innNodes/b_swish_infRV_diff.RData")
bins <- as.character(c(4,6,8,12,16))
detNodesInfBins <- vector("list", length(bins))
names(detNodesInfBins) <- bins

dfInf <- data.frame(Method=character(), Metric=character(), FDR_0.01 = numeric(), FDR_0.05 = numeric(), FDR_0.10 = numeric())
for(i in seq_along(detNodesInfBins))
{
    detNodesInfBins[[bins[i]]] <- sapply(bSwishInf[[i]], function(sw) sw$output[sw$output$signal.node, "node"])
    stats <-  sapply(detNodesInfBins[[bins[i]]], function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
    colnames(stats) <- colnames(dfInf)[c(3:5)]
    stats <- cbind(Method=rep(bins[i],2), Metric = c("FDR", "TPR"), stats)
    dfInf <- rbind(dfInf, stats)
    
}
print(dfInf)
```

##### treeclimbR using levels for stratification
```{r}
load("environment/innNodes/b_swish_levels.RData")
detLevels <- sapply(bSwishLev, function(sw) sw$output[sw$output$signal.node, "node"])
stats <-  sapply(detLevels, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

So treeClimbR on swish has a better control over FDR and higher TPR. Stratifying by inferential variance (Swish), increases both TPR and FDR. treeClimbR on deseq2 has the highest TPR and FDR. This trend is irresepctive of theoretical FDR. Further changing the cuts by infRV does not seem to change things much. The cut 4 seemed to do best in terms of both FDR and TPR. Perhaps decreasing the cuts to 2,3 might be a good check. Level stratification does not seem hold any advantage over infRV. 