---
title: "treeClimbR Results on whole txp"
author: "Noor Pratap Singh"
date: "6/14/2021"
output: html_document
---

```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
library(phangorn)
load("../mikelove-swimdown-216a1dd/simulate/data/simulate.rda")
load("environment/whole_txp/swish_y.RData")
load("environment/whole_txp/tree.RData")
missingTxps <- setdiff(rownames(y), rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[rownames(y),] ###Arranging in the manner of y

innNodes <- nrow(y)+1:tree$Nnode

aggCountsNodes <- computeAggNodes(tree, c(1:nrow(y),innNodes), sim.counts.mat)
fold_changes <- rbind(fold_changes, matrix(0, nrow = length(missingTxps), ncol = ncol(fold_changes),
                                               dimnames = list(missingTxps, colnames(fold_changes))))
logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]/aggCountsNodes[,1]))
fold_changes <- fold_changes[rownames(y),]

trueCountNodes <- which(abs(logFCNodes) >= 0.066) ##Making sure root is non differentiated
trueDriverNodes <-  findDriverGround(trueCountNodes, tree, logFCNodes)[[1]] ### All underlying children have same sign
tLeaf <- which(abs(logFCNodes)[1:nrow(y)] >= 0.066)
print(length(trueDriverNodes))
print(length(tLeaf))
```

#### Swish treeClimbR no stratification

##### Taking driver node leaves as true txps
```{r}
load("environment/whole_txp/bSwish.RData")
detNodes <- sapply(bSwish, function(sw) sw$output[sw$output$signal.node, "node"])
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

##### Taking driver node as ground truth aka node level
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, logFCNodes, tree, type="all"))
print(stats)
```

##### Taking only txps with fold change > 0.066
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

#### Swish treeClimbR stratification with infRV

##### Taking driver node leaves as true txps
```{r}
load("environment/whole_txp/bSwishInfRV.RData")
detNodes <- sapply(bSwishInf, function(sw) sw$output[sw$output$signal.node, "node"])
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

##### Taking driver node as ground truth aka node level
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, logFCNodes, tree, type="all"))
print(stats)
```

##### Taking only txps with fold change > 0.066
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```



#### Swish on leaves

##### Taking driver node leaves as true txps
```{r}
load("environment/whole_txp/ySwishAllNodes.RData")
detNodes <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(y)[,"qvalue"] <= x))
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, pTab=F, type="leaf"))
print(stats)
```

##### Taking only txps with fold change > 0.066
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, pTab=F, type="leaf"))
print(stats)
```

#### Swish on nodes
```{r}
detNodes <- lapply(c(0.01,0.05,0.1), function(x) which(mcols(yAll)[,"qvalue"] <= x))
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, yAll, logFCNodes, tree, pTab=F, type="all"))
print(stats)
```

#### Deseq2 on All

##### Taking driver node leaves as true txps
```{r}
load("environment/whole_txp/bDESeq2.RData")
detNodes <- sapply(bDeseq2, function(de) de$output[de$output$signal.node, "node"])
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

##### Taking driver node as ground truth aka node level
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, logFCNodes, tree, type="all"))
print(stats)
```

##### Taking only txps with fold change > 0.066
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```


#### Deseq2 on nodes stratified by infRV

##### Taking driver node leaves as true txps
```{r}
load("environment/whole_txp/b_deseq2_infRV.RData")
detNodes <- sapply(bDeseq2Inf, function(de) de$output[de$output$signal.node, "node"])
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, pTab=T, type="leaf"))
print(stats)
```

##### Taking driver node as ground truth aka node level
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, logFCNodes, tree, type="all"))
print(stats)
```

##### Taking only txps with fold change > 0.066
```{r}
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

#### Deseq2 on leaves
```{r}
load("environment/whole_txp/res.RData")
detNodes <- lapply(c(0.01,0.05,0.1), function(x) which(res$padj[1:nrow(y)] <= x))
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, y, NULL, tree, type="leaf"))
print(stats)
stats <-  sapply(detNodes, function(nodes) computeTPFP(tLeaf, nodes, y, NULL, tree, type="leaf"))
print(stats)
```

#### Deseq2 on nodes
```{r}
detNodes <- lapply(c(0.01,0.05,0.1), function(x) which(res$padj <= x))
stats <-  sapply(detNodes, function(nodes) computeTPFP(trueDriverNodes, nodes, res, logFCNodes, tree, pTab=F, type="all"))
print(stats)
```