---
title: "explore fns"
author: "Noor Pratap Singh"
date: "4/10/2022"
output: html_document
---

Trying to see what transcripts/genes can be recovered from tree climbing
1) Ideally has atleast one significant ancestor (defined by qvalue)
2) no significant sibling
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/fs/cbcb-lab/rob/students/noor/Uncertainity/treeTerminusScripts')
```

```{r}
load("environment/brain_sim_nodtu_lowfc/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/yAll.RData")
load("environment/brain_sim_nodtu_lowfc/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/y.RData")
load("environment/brain_sim_nodtu_lowfc/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/tree.RData")
load("environment/brain_sim_nodtu_lowfc/simulate.rda")
load("environment/brain_sim_nodtu/mode=gc_bias/mapDf.RData")

suppressPackageStartupMessages(source("tree_helper_function.R"))
suppressPackageStartupMessages(source("tree_term_climb.R"))
signs <- computeSign(yAll, "condition", minP=0.80)
missingTxps <- setdiff(tree$tip, rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[tree$tip,]
length(missingTxps)

innNodes <- length(tree$tip)+1:tree$Nnode
aggCountsNodes <- computeAggNodesU(tree, c(1:length(tree$tip),innNodes), sim.counts.mat)

logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]+1)-log2(aggCountsNodes[,1]+1))
iso.any <- iso.dtu | iso.dte | iso.dge
dts <- intersect(names(iso.any)[iso.any], names(logFCNodes))

dtInds <- match(dts, rownames(y))
sTxps <- which(mcols(y)[["qvalue"]] < 0.1)

trueG <- union(dge.genes, dte.genes)
sGenes <- unique(mapDf[rownames(y)[sTxps],1])
print(length(sGenes))
print(length(intersect(sGenes, trueG)))
```
We have a total of 2603 genes identified by us out of which 2315are true dges.
1835 genes are still missed.
We look at the transcripts of these 1835 genes and see how many of them can be retrieved by the tree structure
These map to 10813 txps out which 713 map to Ancestors so cant be salavaged
Only 289 txps have ancestors that have ancestors meeting lfdr pvalue
```{r}
fGenes <- setdiff(trueG, sGenes)
print(length(fGenes))
fGTXps <- sapply(fGenes, function(g) match(mapDf[,2][mapDf[,1] == g], rownames(y)))

fT <- unlist(fGTXps)
fT <- fT[!is.na(fT)]

fTxps <- fT

Anc <- Ancestors(tree, fTxps, "parent")
print(sum(Anc == nrow(y)+1)) ### Txps that have root as parent aka cant be saved by tree

fTxps <- fTxps[Anc != nrow(y)+1]
Ancs <- Ancestors(tree, fTxps)

###Finding lfdr threshold
library(fdrtool)
pvalsInn <- mcols(yAll)[["pvalue"]][nrow(y)+1:tree$Nnode]
naInds <- is.na(pvalsInn)
fdrIn <- fdrtool(pvalsInn[!naInds], statistic="pvalue")
pvalAdjIn <- rep(1, length(naInds))
pvalAdjIn[!naInds] <- fdrIn[["lfdr"]]
plfdr <- pvalsInn[which.min(abs(pvalAdjIn - 0.1))] ##0.1 thresh
print(plfdr)
sig <- sapply(Ancs, function(anc) sum(mcols(yAll)[anc,"pvalue"] < plfdr, na.rm=T) > 0) #474 txps that have atleast one sig ancestor
fTxps <- fTxps[sig]
print(length(fTxps))
Ancs <- Ancs[sig]
AncsM <- sapply(Ancs, function(anc) max(anc[which(mcols(yAll)[anc,"pvalue"] < plfdr)]))

sameSigns <- sapply(Descendants(tree, AncsM, "all"), function(nodes) all(signs[nodes]>=0) | all(signs[nodes]<=0)) ##Out of these 356 txps have ancestors with same signs
print(sum(sameSigns)) 

rTxps <- fTxps[sameSigns]
Ancs <- Ancs[sameSigns]
AncsM <- AncsM[sameSigns]

unSTxps <- setdiff(unlist(fT), rTxps) ##Transcripts that cant be saved
unG <- unique(mapDf[rownames(y)[unSTxps],1]) #1742 genes can't be saved
print(length(unG))

rG <- unique(mapDf[rownames(y)[rTxps],1]) #57 genes can be saved so far
print(length(rTxps))
print(length(rG))

sibs <- Siblings(tree, rTxps) ###Siblings
sSib <- sapply(sibs, function(x) {
  xIn <- x[x > nrow(y)]
  xL <- x[x <= nrow(y)]
  inn <- F
  leaf <- F
  if(length(xL) > 0 & sum(mcols(yAll)[xL, "qvalue"] < 0.1, na.rm = T) > 0)
      leaf <- T
  if(length(xIn) > 0 & sum(mcols(yAll)[xIn, "pvalue"] < 0.0012, na.rm = T) > 0)
      inn <- T
  return(inn | leaf)
})
print(sum(sSib)) ## 54 txps have siblings that are significant
rTxps <- rTxps[!sSib]
Ancs <- Ancs[!sSib]
AncsM <- AncsM[!sSib]

mirvI <- sapply(Descendants(tree, AncsM), function(nodes) sum(mcols(yAll)[nodes, "meanInfRV"] < 0.2) > 0) ## 145 have txps that dont meet infRV criteria
print(sum(mirvI))
rTxps <- rTxps[!mirvI]
Ancs <- Ancs[!mirvI]
print(length(rTxps))
rGe <- unique(mapDf[rownames(y)[rTxps],1]) 
print(length(rGe)) ##44 Genes
```

#### Using a higher pvalue threshold
```{r}
fTxps <- fT

Anc <- Ancestors(tree, fTxps, "parent")
print(sum(Anc == nrow(y)+1)) ### Txps that have root as parent aka cant be saved by tree

fTxps <- fTxps[Anc != nrow(y)+1]
Ancs <- Ancestors(tree, fTxps)
pThr <- estimatePThresh(yAll[1:nrow(y),],0.1)
print(pThr)
sig <- sapply(Ancs, function(anc) sum(mcols(yAll)[anc,"pvalue"] < pThr, na.rm=T) > 0) #842 txps that have atleast one sig ancestor
fTxps <- fTxps[sig]
print(length(fTxps))
Ancs <- Ancs[sig]
ls <- sapply(Ancs,length)
AncsM <- sapply(Ancs, function(anc) max(anc[which(mcols(yAll)[anc,"pvalue"] < pThr)]))

sameSigns <- sapply(Descendants(tree, AncsM, "all"), function(nodes) all(signs[nodes]>=0) | all(signs[nodes]<=0)) ##Out of these 689 txps have ancestors with same signs
print(sum(sameSigns))

rTxps <- fTxps[sameSigns]
Ancs <- Ancs[sameSigns]
AncsM <- AncsM[sameSigns]

unSTxps <- setdiff(unlist(fT), rTxps) ##Transcripts that cant be saved
unG <- unique(mapDf[rownames(y)[unSTxps],1]) #1708 genes can't be saved
print(length(unG))

rG <- unique(mapDf[rownames(y)[rTxps],1]) #118 genes can be saved so far
print(length(rTxps))
print(length(rG))

sibs <- Siblings(tree, rTxps) ###Siblings
sSib <- sapply(sibs, function(x) {
  xIn <- x[x > nrow(y)]
  xL <- x[x <= nrow(y)]
  inn <- F
  leaf <- F
  if(length(xL) > 0 & sum(mcols(yAll)[xL, "qvalue"] < 0.1, na.rm = T) > 0)
      leaf <- T
  if(length(xIn) > 0 & sum(mcols(yAll)[xIn, "pvalue"] < 0.0012, na.rm = T) > 0)
      inn <- T
  return(inn | leaf)
})
print(sum(sSib)) ## 59 txps have siblings that are significant
rTxps <- rTxps[!sSib]
Ancs <- Ancs[!sSib]
AncsM <- AncsM[!sSib]

mirvI <- sapply(Descendants(tree, AncsM), function(nodes) sum(mcols(yAll)[nodes, "meanInfRV"] < 0.2) > 0) ## 252 dont meet the infrv criteria
print(sum(mirvI))
rTxps <- rTxps[!mirvI]
Ancs <- Ancs[!mirvI]
print(length(rTxps))
rGe <- unique(mapDf[rownames(y)[rTxps],1]) ##98 genes more
print(length(rGe)) 
```