---
title: "Chimpanzee Bulk Data"
author: "Noor Pratap Singh"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/fs/cbcb-lab/rob/students/noor/Uncertainity/treeTerminusScripts')
```

#### Loading Data
```{r}
suppressPackageStartupMessages(library(tximeta))
suppressPackageStartupMessages(library(SummarizedExperiment))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(fishpond))
metaData <- read.delim("/fs/cbcb-lab/rob/students/noor/Uncertainity/ChimpData/SYNAPSE_METADATA_MANIFEST.tsv")
quantDir <- "/fs/cbcb-lab/rob/students/noor/Uncertainity/ChimpData/out_sal"
samples <- sapply(strsplit(metaData[,"path"], split = "/", fixed = T), function(x) x[[length(x)]])
samples <- sapply(strsplit(samples, ".", fixed = T), function(x) x[[1]])
files <- file.path(quantDir, samples, "quant.sf")
colData <- cbind(data.frame(files = files, names = samples), metaData)
colData[["condition"]] <- ifelse(colData[["tissue"]] == "cerebellum", "Cere", "Rest")
# seChimp <- tximeta::tximeta(colData)
# save(seChimp, file = "environment/Chimp/se_chimp.RData")
load("environment/Chimp/se_chimp.RData")
colData(seChimp)[["condition"]] <- as.factor(colData[["condition"]])
```

### Some statistics
```{r}
print(table(metaData[,"tissue"])) ### Number of replicates per tissue
print(table(metaData[,"individualID"])) ### Number of replicates per patient
hist(colSums(assays(seChimp)[["counts"]]))
df <- data.frame(depth=colSums(assays(seChimp)[["counts"]]), patient = metaData[,"individualID"])
print(df %>% group_by(patient) %>% summarise(mDepth = mean(depth)))

print(sum(rowSums(assays(seChimp)[["counts"]]) <= 10))
print(sum(apply(assays(seChimp)[["counts"]], 1, function(x) sum(x != 0) == 4)))
```


#### PCA
```{r}
sampDorNucl <- colData %>% filter(tissue == "medial dorsal nucleus of thalamus") %>% select(names) %>% unlist
dds <- DESeqDataSet(seChimp, design=~tissue)
dds <- estimateSizeFactors(dds)
dds <- dds[rowSums(assays(seChimp)[["counts"]]) > 10,]
vsd <- vst(dds)

plotPCA(vsd, intgroup=c("tissue"), ntop=1000)
plotPCA(vsd, intgroup=c("individualID"), ntop=1000)

plotPCA(vsd, intgroup=c("tissue"), ntop=500)
```
There seems to exist a difference between cerebellum and medial dorsal nucleus. However, the patient clearly seems different and has the lowest depth.

#### Removing samples from patient 162
```{r}
samp162 <- colData %>% filter(individualID == "PTB162") %>% select(names) %>% unlist
vsdR162 <- vst(dds[,!(colnames(dds) %in% samp162)])
plotPCA(vsdR162, intgroup=c("tissue"), ntop=1000)
plotPCA(vsdR162, intgroup=c("individualID"), ntop=1000)
rm(vsdR162, vsd)
gc()
```
There seems batch effects now due to sample 165 well

<!-- #### Removing samples from medial prefrontal cortex and 162 -->
<!-- ```{r} -->

<!-- vsdRDorNucl <- vst(dds[,!(colnames(dds) %in% sampDorNucl)]) -->
<!-- plotPCA(vsdRDorNucl, intgroup=c("tissue"), ntop=1000) -->
<!-- plotPCA(vsdRDorNucl, intgroup=c("individualID"), ntop=1000) -->
<!-- rm(vsdRDorNucl, vsdR162, vsd) -->
<!-- gc() -->
<!-- ``` -->
<!-- Now we see sample 165 as separated  -->

#### SVA on entire dataset removing dorsal samples
```{r}
dds <- dds[,!(colnames(dds) %in% sampDorNucl)]
print(dim(dds))
suppressPackageStartupMessages(library(sva))
suppressPackageStartupMessages(library(ggplot2))
mm <- model.matrix(~ condition, colData(dds))
mm0 <- model.matrix(~ 1, colData(dds))

norm.cts <- counts(dds, normalized=TRUE)
norm.cts <- norm.cts[rowSums(norm.cts) > 0,]
fit <- svaseq(norm.cts, mod=mm, mod0=mm0, n.sv=2)
df <- data.frame(SV1=fit$sv[,1], SV2=fit$sv[,2], cond = colData(dds)[,"condition"], indID = colData(dds)[,"individualID"])
ggplot(df, aes(SV1, SV2, color=indID)) + geom_point()
rm(norm.cts)
gc()
```
We see 162,165 clearly shown separately by SV

#### SVA without removing 162
```{r}
ddsR162 <- dds[,!(colnames(dds) %in% samp162)]
norm.cts <- counts(ddsR162, normalized=TRUE)
mm <- model.matrix(~ condition, colData(ddsR162))
mm0 <- model.matrix(~ 1, colData(ddsR162))
norm.cts <- norm.cts[rowSums(norm.cts) > 0,]
fitR162 <- svaseq(norm.cts, mod=mm, mod0=mm0, n.sv=2)
df <- data.frame(SV1=fitR162$sv[,1], SV2=fitR162$sv[,2], cond = colData(ddsR162)[,"condition"], indID = colData(ddsR162)[,"individualID"])
ggplot(df, aes(SV1, SV2, color=indID)) + geom_point()
rm(dds, ddsR162, norm.cts)
gc()
```
All samples seem to be captured here by SVA

#### Batch Effect correction
```{r}
corrBESwish <- function(se, samples, fit)
{
    y <- se[,!(colnames(se) %in% samples)]
    print(dim(y))
    y <- scaleInfReps(y, saveMeanScaled=TRUE)
    y <- labelKeep(y)
    y <- y[mcols(y)$keep,]
    print(dim(y))
    
    infRepIdx <- grep("infRep",assayNames(y),value=TRUE)
    nreps <- length(infRepIdx)
    
    mm <- model.matrix(~condition, colData(y))
    
    pc <- .1
    for (k in seq_len(nreps)) {
      logInfRep <- log(assay(y, infRepIdx[k]) + pc)
      logInfRep <- limma::removeBatchEffect(
                            logInfRep,
                            covariates=fit[["sv"]],
                            design=mm)
      assay(y, infRepIdx[k]) <- exp(logInfRep)
    }
    gc()
    y <- swish(y, x="condition")
    y
}
```

#### Running Swish after batch correction
```{r}
# y <- corrBESwish(seChimp, sampDorNucl, fit) ## Swish on all samples
# save(y, file = "environment/Chimp/y.RData")
load("environment/Chimp/y.RData")
dd <- mcols(y)
dd <- dd[order(dd[["qvalue"]]),]
rm(y)
gc()

# yR162 <- corrBESwish(seChimp, union(sampDorNucl, samp162), fitR162) ## Swish by removing 162
# save(yR162, file = "environment/Chimp/yR162.RData")
load("environment/Chimp/yR162.RData")
ddR162 <- mcols(yR162)
ddR162 <- ddR162[order(ddR162[["qvalue"]]),]
rm(yR162)
gc()
```

#### Total signficant txps and the intersection of the top z txps based on qvalues
```{r}
print(sum(dd[["qvalue"]] < 0.05 & abs(dd[["log2FC"]]) > 1))
print(sum(ddR162[["qvalue"]] < 0.05 & abs(ddR162[["log2FC"]])>1))

print(sapply(c(10,100,500,1000,2000,5000,7000,10000), function(x) {
  length(intersect(rownames(dd)[1:x], rownames(ddR162)[1:x]))
}))
```
There exists a lot of intersection

```{r fig.width=12, fig.height=12}
load("environment/Chimp/y.RData")

hist(mcols(y)$pvalue)
sig <- mcols(y)$qvalue < .05
lo <- order(mcols(y)$log2FC * sig)
hi <- order(-mcols(y)$log2FC * sig)

with(mcols(y),
     table(sig=qvalue < .05, sign.lfc=sign(log2FC))
     )
top_up <- mcols(y)[head(hi),]
print(top_up)
top_down <- mcols(y)[head(lo),]
print(top_down)

pL <- list()
plotMASwish(y, alpha = 0.05)
par(mfrow=c(2,5))
for(i in c(1,2,6,7,30,48,59,86,110,211))
  plotInfReps(y, idx=hi[i], x="condition")

par(mfrow=c(2,5))
for(i in c(257,384,471,619,736,764,800,900,950,1000))
  plotInfReps(y, idx=hi[i], x="condition")

par(mfrow=c(2,5))
for(i in c(1,29,34,37,59,61,154,163,201,250))
  plotInfReps(y, idx=lo[i], x="condition")
par(mfrow=c(1,5))
for(i in c(300,500,600,700,800))
  plotInfReps(y, idx=lo[i], x="condition")

```
