---
title: "Read Trees"
author: "Noor Pratap Singh"
date: "8/18/2021"
output: html_document
---

This file reads the tree terminus output of running on Chimpanzee bulk data and preparing the bulk and tree output accordingly
Using output from read_sal, where we have seen the existence of batch effects and we also removed medial dorsal nucleus of thalmus samples for the analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/cbcb_rob/Uncertainity/treeTerminusScripts')
```

#### Reading the nwk tree
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(sva))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
source("tree_helper_function.R")
load("environment/Chimp/se_chimp.RData")
trees <- read.tree("/fs/cbcb-lab/rob/students/noor/Uncertainity/ChimpData/terminus/cluster_nwk.txt")
tLengths <- sapply(trees, function(x) length(x$tip.label))
print(length(tLengths))
print(max(tLengths))
print(sum(tLengths >= 10))
tree <- mergeTree(trees, se = seChimp)
print(tree)
save(tree, file="environment/Chimp/tree.RData")
```
Only 14 groups with total number of txps >= 10. Max group length is 14 with total number of groups 5794. A total of 15420 txps covered by tree with 9435 internal nodes

#### Processing the data running swish
```{r}
# sampDorNucl <- data.frame(colData(seChimp)) %>% filter(tissue == "medial dorsal nucleus of thalamus") %>% dplyr::select(names) %>% unlist
# y <- labelKeep(seChimp[,!(colnames(seChimp) %in% sampDorNucl)])
# tnames <- union(tree$tip.label, rownames(y)[mcols(y)$keep])
# y <- y[tnames,]
# mcols(y)[,'keep'] <- TRUE
# dds <- DESeqDataSet(y, design=~condition)
# dds <- estimateSizeFactors(dds)
# mm <- model.matrix(~ condition, colData(dds))
# mm0 <- model.matrix(~ 1, colData(dds))
# 
# norm.cts <- counts(dds, normalized=TRUE)
# norm.cts <- norm.cts[rowSums(norm.cts) > 0,]
# fit <- svaseq(norm.cts, mod=mm, mod0=mm0, n.sv=2)
# df <- data.frame(SV1=fit$sv[,1], SV2=fit$sv[,2], cond = colData(dds)[,"condition"], indID = colData(dds)[,"individualID"])
# ggplot(df, aes(SV1, SV2, color=indID)) + geom_point()
# 
# # y <- corrBESwish(y, sampDorNucl, fit) ### Correcting for batch effects
# yU <- mergeLeaves(tree, y)
# tree <- yU[[1]]
# y <- yU[[2]]
# rm(yU)
# gc()
# hist(mcols(yAll)[["pvalue"]])
# y <- computeInfRV(y)
# save(y, file = "environment/Chimp/yrem_swish.RData")
# save(tree, file="environment/Chimp/tree.RData")
```

#### Preparing yAll
```{r}
# yAll <- prepSwish(tree, y)
# yAll <- scaleInfReps(yAll, lengthCorrect = T)
# yAll <- labelKeep(yAll)
# mcols(yAll)$keep=T
# yAll <- swish(yAll, x = "condition")
# yAll <- computeInfRV(yAll)
# hist(mcols(yAll)[["pvalue"]])
# save(yAll, file = "environment/Chimp/yAll.RData")
```