---
title: "R Notebook"
output: html_notebook
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/fs/cbcb-lab/rob/students/noor/Uncertainity/treeTerminusScripts')
```

```{r}
suppressPackageStartupMessages(source("tree_helper_function.R"))
suppressPackageStartupMessages(source("tree_term_climb.R"))
load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/yAll.RData")
load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/y.RData")
load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/tree.RData")
load("environment/brain_sim_nodtu/simulate.rda")

load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/tAfterQvalue.RData")
load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/tAfterQvalueNoStop.RData")
load("environment/brain_sim_nodtu/mode=gc_bias/post_type=gibbs_nrep=100_tf=100/tAfterQvalue_NoSign.RData")
```

```{r}
signs <- computeSign(yAll, "condition")
missingTxps <- setdiff(tree$tip, rownames(sim.counts.mat))
sim.counts.mat <- rbind(sim.counts.mat, matrix(0, nrow = length(missingTxps), ncol = ncol(sim.counts.mat),
                                               dimnames = list(missingTxps, colnames(sim.counts.mat))))
sim.counts.mat <- sim.counts.mat[tree$tip,]
length(missingTxps)

innNodes <- length(tree$tip)+1:tree$Nnode
aggCountsNodes <- computeAggNodesU(tree, c(1:length(tree$tip),innNodes), sim.counts.mat)

logFCNodes <- ifelse(rowSums(aggCountsNodes)==0, 0, log2(aggCountsNodes[,2]+1)-log2(aggCountsNodes[,1]+1))
iso.any <- iso.dtu | iso.dte | iso.dge
dts <- intersect(names(iso.any)[iso.any], names(logFCNodes))
rootFC <- abs(logFCNodes[length(tree$tip)+1])
dts <- setdiff(dts, dts[which(abs(logFCNodes[dts]) <= rootFC)])

print(mean(abs(logFCNodes[dts]) > rootFC))
```


```{r}
detNodes <- list()
detNodes[["Swish"]] <- lapply(c(0.01, 0.05, 0.1), function(x) which(mcols(y)[,"qvalue"] <= x ))
detNodes[["tAfterqval"]] <- lapply(tAfterQvalue, function(ta) ta[["candNodeO"]])
detNodes[["tAfterqvalNoSign"]] <- lapply(tAfterQvalueNoSign, function(ta) ta[["candNodeO"]])
detNodes[["tAfterqvalNoStop"]] <- lapply(tAfterQvalueNoStop, function(ta) ta[["candNodeO"]])

inds <- lapply(tAfterQvalue, function(ta) ta[["resDf"]][,"inds"])
detNodes[["tAfterqvalAllNodes"]] <- lapply(seq(3), function(i) inds[[i]][mcols(yAll)[inds[[i]],"qvalue"] <= c(0.01,0.05,0.1)[i]])

nodes <- lapply(tAfterQvalue, function(ta) c(ta[["naNodeO"]], ta[["candNodeO"]], ta[["negNodeO"]]))
qvalues <- lapply(nodes, function(node) p.adjust(mcols(yAll)[node,"pvalue"], method="BH"))
detNodes[["tAfterqvalOnlyNodes"]] <- lapply(seq_along(qvalues), function(i) nodes[[i]][which(qvalues[[i]] <= c(0.01,0.05,0.1)[i])])
```

```{r}
innNodesE <- detNodes[["tAfterqval"]][[3]][detNodes[["tAfterqval"]][[3]] > nrow(y)]
posNodes <- innNodesE[which(abs(logFCNodes)[innNodesE] > rootFC)]
table(sapply(posNodes,length))
gDiff <- setdiff(unique(rowData(y)[,1]), c(dge.genes, dte.genes))
innG <- sapply(Descendants(tree,posNodes), function(nodes) mapDf[rownames(y)[nodes],1])
table(sapply(innG, length)) ## Number of descendant leaves in each node
table(sapply(innG, function(g) length(unique(g)))) ### Number of genes associated with each node
print(length(unique(unlist(innG)))) ### Total number of unique genes
print(sum(unique(unlist(innG)) %in% dte.genes)) ### unique genes in dte
print(sum(unique(unlist(innG)) %in% dge.genes))
missG <- unique(unlist(innG))[unique(unlist(innG)) %in% gDiff] ## genes with inner nodes that neither dte or dge
which(sapply(innG, function(x) sum(x %in% missG)>0))

test <- abs(mcols(yAll)[innNodesE, "stat"])
fp <- rep(1, length(test))
fp[which(abs(logFCNodes)[innNodesE] < rootFC)] = 2
fp <- fp[order(test, decreasing = T)]
test <- sort(test, decreasing = T)
plot(test, col=fp, cex=1.5, pch=20)

innNodesE <- detNodes[["tAfterqvalOnlyNodes"]][[3]][detNodes[["tAfterqvalOnlyNodes"]][[3]] > nrow(y)]
test <- abs(mcols(yAll)[innNodesE, "stat"])
fp <- rep(1, length(test))
fp[which(abs(logFCNodes)[innNodesE] < rootFC)] = 2
fp <- fp[order(test, decreasing = T)]
test <- sort(test, decreasing = T)
plot(test, col=fp, cex=1.5, pch=20)
```

```{r}
negNodes <- list()
negNodes[["Swish"]] <- lapply(detNodes[["Swish"]], function(x) setdiff(seq(nrow(y)), x))
negNodes[["tAfterqval"]] <- lapply(tAfterQvalue, function(ta) unlist(Descendants(tree,c(ta[["naNodeO"]],ta[["negNodeO"]]))))
negNodes[["tAfterqvalNoSign"]] <- lapply(tAfterQvalueNoSign, function(ta) unlist(Descendants(tree,c(ta[["naNodeO"]],ta[["negNodeO"]]))))
negNodes[["tAfterqvalNoStop"]] <- lapply(tAfterQvalueNoStop, function(ta) unlist(Descendants(tree,c(ta[["naNodeO"]],ta[["negNodeO"]]))))
negNodes[["tAfterqvalAllNodes"]] <- lapply(detNodes[["tAfterqvalAllNodes"]], function(nodes) setdiff(seq(nrow(y)), unlist(Descendants(tree, nodes))))
negNodes[["tAfterqvalOnlyNodes"]] <- lapply(detNodes[["tAfterqvalOnlyNodes"]], function(nodes) setdiff(seq(nrow(y)), unlist(Descendants(tree, nodes))))

df <- data.frame(Method=character(), Metric=character(), FDR_0.01 = numeric(), FDR_0.05 = numeric(), FDR_0.10 = numeric())
for(n in names(detNodes))
{
    print(n)
    stats <-  sapply(seq_along(detNodes[[n]]), function(i) computeMetOut(detNodes[[n]][[i]], logFCNodes, tree = tree, negNodes = negNodes[[n]][[i]], lfcThresh = rootFC-0.001))
    colnames(stats) <- colnames(df)[c(3:5)]
    stats <- cbind(Method=rep(n,2), Metric = c("FDR", "TPR"), stats)
    df <- rbind(df, stats)
}
print(df)

```

```{r}
print(sum(detNodes[[2]][[3]] > nrow(y)))
innNodes <- detNodes[[2]][[3]][detNodes[[2]][[3]] > nrow(y)]
sum(abs(logFCNodes[innNodes]) < rootFC)
```
A total of 230 inner nodes detected, out of which 78 are false positives

```{r}
ngNodes <- c(tAfterQvalue[[3]][["naNodeO"]], tAfterQvalue[[3]][["negNodeO"]])
ngNodes <- ngNodes[order(mcols(yAll)[ngNodes, 'meanInfRV'], decreasing = T)]
tpNodes <- match(names(logFCNodes)[abs(logFCNodes) > rootFC], rownames(yAll))
tnNodes <- match(names(logFCNodes)[abs(logFCNodes) < rootFC], rownames(yAll))
tns <- intersect(tnNodes, unlist(Descendants(tree,ngNodes)))
fns <- intersect(tpNodes, ngNodes)
fns <- fns[order(mcols(yAll)[fns, 'meanInfRV'], decreasing = T)]
fnsL <- fns[fns < nrow(y)]
print(length(fnsL))

pmatch <- fnsL[which(mcols(yAll)[["pvalue"]][fnsL] < tAfterQvalue[[3]][["pChild"]])] ##All leaves that have pvalue less than threshold
sum(tAfterQvalue[[3]][["resDf"]][match(pmatch, tAfterQvalue[[3]][["resDf"]][,3]),2] > 0.1)
```
141 missed due to multi hypothesis correction

```{r}
fnsL <- setdiff(fnsL, pmatch)
lNodes <- fnsL[sapply(Ancestors(tree, fnsL), function(x) x[1] == (nrow(y)+1))]
print(length(lNodes))
```
388 transcripts have root as their direct ancestor

```{r}
fnsL <- setdiff(fnsL, lNodes)
oppNodes <- fnsL[sapply(fnsL, function(node) {
  sig <- signs[Descendants(tree, Ancestors(tree, node)[1])[[1]]]
  if(all(sig >= 0, na.rm=T) | all(sig <= 0, na.rm=T))
      return(F)
  return(T)
})]
print(length(oppNodes))
anc <- Ancestors(tree, oppNodes)
oppAnc <- oppNodes[sapply(anc, function(nodes) sum(mcols(yAll)[nodes,"pvalue"] < tAfterQvalue[[3]][["pChild"]]) > 0)]
print(length(oppAnc))
SibO <- Siblings(tree, oppNodes)

i <- 1
print(mcols(yAll)[c(oppNodes[i],Descendants(tree,SibO[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,SibO[[i]])[1],])

i <- 10
print(mcols(yAll)[c(oppNodes[i],Descendants(tree,SibO[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,oppNodes[[i]])[1],])

i <- 100
print(mcols(yAll)[c(oppNodes[i],Descendants(tree,SibO[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,oppNodes[i])[1],])


```
345 transcripts that had siblings with opposite signs, out of these 94 have ancestors that have significant pvalues

```{r}
library(ggplot2)
library(ggpubr)

i <- 1
df1 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 1)
df2 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 2)

p1 <- df1 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
p2 <- df2 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
ggarrange(p1, p2, common.legend=T)

i <- 10
df1 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 1)
df2 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 2)

p1 <- df1 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
p2 <- df2 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
ggarrange(p1, p2, common.legend=T)

i <- 100
df1 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 1)
df2 <- getTrace(yAll, c(oppNodes[i], SibO[[i]], Ancestors(tree, oppNodes[i])[1]), 2)

p1 <- df1 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
p2 <- df2 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
ggarrange(p1, p2, common.legend=T)

i <- 1
df1 <- getTrace(yAll, c(oppAnc[i], Siblings(tree, oppAnc[i])[[1]], Ancestors(tree, oppAnc[i])[1]), 1)
df2 <- getTrace(yAll, c(oppAnc[i], Siblings(tree, oppAnc[i])[[1]], Ancestors(tree, oppAnc[i])[1]), 2)

p1 <- df1 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
p2 <- df2 %>% ggplot(aes(infRep, count, col=factor(feature))) + geom_point() + geom_line()
ggarrange(p1, p2, common.legend=T)
```

```{r}
fnsL <- setdiff(fnsL, oppNodes)
Sibs <- Siblings(tree, fnsL)
sibFN <- fnsL[which(sapply(Sibs, function(sib) sum(mcols(yAll)[c(sib,unlist(Descendants(tree,sib,"all"))),"pvalue"] < tAfterQvalue[[3]][["pChild"]], na.rm=T) > 0))]
Sibs <- Sibs[match(sibFN, fnsL)]
print(length(sibFN))

i <- 1
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 10
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 20
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 30
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 100
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 200
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 400
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 600
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])

i <- 800
print(mcols(yAll)[c(sibFN[i],Descendants(tree,Sibs[[i]])[[1]]),])
print(mcols(yAll)[Ancestors(tree,Sibs[[i]])[1],])
```
919 have siblings or either their descendants that have significant pvalue

```{r}
fnsL <- setdiff(fnsL, sibFN)
fnInf <- fnsL[sapply(fnsL, function(node) sum(mcols(yAll)[c(Siblings(tree, node), node),"meanInfRV"] < 0.3) > 0)]
fnsL <- setdiff(fnsL, fnInf)
```
215 nodes dont climb because of infRV criteria

For the remaining 3 nodes, descendants of siblings had already been looked

Look at ancestor nodes for oppNodes